Project Proposal: Implementing Flow-based Balanced Bipartition Algorithm

### 2014-04-25 sequential/latex/abstraction/prelim_gf.tex
upload files from laptop linux

A Galois field $\Fq$ is a field with a finite number of elements. The
number of elements $q$ of the field is a power of a prime integer ---
i.e. $q = p^k$, where $p$ is a prime integer, and $k \geq 1$
\cite{galois_field:mceliece}. 
%Galois fields are denoted as ${\mathbb{F}}_{q}$ and also $GF(q
%=p^k)$. 
We are interested in fields where $p = 2$ and $k >1$ --- i.e. {\it
  binary Galois extension   fields} $\mathbb{F}_{2^k}$ --- as they are
widely employed in hardware implementations of cryptography
primitives.

To construct ${\mathbb{F}}_{2^k}$, we take the polynomial ring
${\mathbb{F}}_2[x]$, where ${\mathbb{F}}_{2} = \{0, 1\}$, and an
irreducible  polynomial $P(x) \in {\mathbb{F}}_2[x]$ of degree $k$, and
construct ${\mathbb{F}}_{2^k}$ as ${\mathbb{F}}_2[x] \pmod{   P(x)}$. As a
result, all field operations are performed 
modulo the irreducible polynomial $P(x)$ and the coefficients are
reduced modulo $p=2$. %=2$; due to which $-1 = +1$ over $\Fkk$. 
Any element $A \in {\mathbb{F}}_{2^k}$ can be represented in polynomial
form as $A = a_0 +  a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}$, where
$a_i \in {\mathbb{F}}_2, i = 0, \dots, k-1$, and $\alpha$ is the root of
the irreducible polynomial, {\it i.e.} $P(\alpha)=0$. 
%The field
%$\Fkk$ can therefore be construed as a $k$-dimensional vector space
%over ${\mathbb{F}}_2$.

@@

\begin{Example}\label{ex:1}
{\it
Let us construct ${\mathbb{F}}_{2^4}$ as ${\mathbb{F}}_2[x] \pmod{
  P(x)}$, where $P(x)=x^4+x^3+1 \in {\mathbb{F}}_2[x]$ is an
irreducible polynomial of degree $k =4$. Let $\alpha$ be the root of
$P(x)$, i.e. $P(\alpha)=0$.  Any element $A \in {\mathbb{F}}_2[x]
\pmod{ x^4 + x^3 + 1}$ has a representation of the type: $A = a_3 x^3
+ a_2 x^2 +  a_1 x + a_0$ where the coefficients $a_3, \dots, a_0$ are
in ${\mathbb{F}}_2 = \{0, 1\}$. Since there are only 16 such
polynomials, we obtain 16 elements in the field
${\mathbb{F}}_{16}$. Each element in can then be viewed as a 4-bit
vector over ${\mathbb{F}}_2$: ${\mathbb{F}}_{16}=\{(0000),(0001),
\dots (1110),(1111)\}$.  Each element also has an exponential
representation; all three representations are shown in Table
\ref{tab:gfelement}. For example, consider the element $\alpha^{12}$.
Computing $\alpha^{12} \pmod{ \alpha^4+\alpha^3+1} = \alpha + 1
= (0011)$; hence we have the three equivalent representations. 
}

\vspace{-0.1in}
\begin{table}[h]
\begin{center}
{\tiny
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
\pmod{x^4+x^3+1}$}\label{tab:gfelement}  
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
}
\end{center}
\end{table}
\end{Example}

%%%%%%%%%
{\bf Polynomial Functions $f: \Fkk \rightarrow \Fkk$:} 
Arbitrary mappings among $k$-bit vectors can be constructed; each such
mapping generates a function $f: \B^k \rightarrow \B^k$. 
%. Since every
%$k$-bit vector can be construed as an element in $\Fkk$ (as shown in
%the above example), every such function corresponds to a function over
Every such function is also a polynomial function over Galois fields:
$f: \Fkk \rightarrow \Fkk$.

\begin{Theorem}
From \cite{ff:1997}: 
Any  function $f: \Fq \to \Fq$ is a polynomial function
over $\Fq$, that is there exists a polynomial $\F \in \Fq [x]$ such
  that $f(a) = \F(a)$, for all $a \in \Fq$. 
\end{Theorem}

@@

%% \begin{Example} {\it
%% Let $A = \{a_2, a_1, a_0\}$ and $Y = \{y_2, y_1,
%% y_0\}$ be 3-bit vectors.  Consider the function $Y[2:0] = A[2:0]
%% >> 1$, i.e. a {\bf bit-vector right shift} operation on $A$. 
%% The function maps as follows:

%% \begin{center}
%% {\small
%% \begin{tabular}{c|ccc|c||c|ccc|c}
%% $\{a_2a_1a_0\}$  & $A$ &$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$ &
%%   $\{a_2a_1a_0\}$  &$A$&$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$\\
%% \hline
%% 000  &0 &$\rightarrow$& 000 & 0 & 100  &$\alpha^2$ &$\rightarrow$& 010 &  $\alpha$ \\
%% 001  &1 &$\rightarrow$& 000 & 0 & 101  &$\alpha^2 + 1$ &$\rightarrow$&010 & $\alpha$ \\
%% 010  &$\alpha$ & $\rightarrow$ & 001& 1 & 110  &$\alpha^2 + \alpha$&$\rightarrow$& 011 &$\alpha + 1$ \\   
%% 011  &$\alpha + 1$ &$\rightarrow$& 001 &1 & 111& $\alpha^2 + \alpha + 1$ &$\rightarrow$& 011 &$\alpha + 1$\\
%% \hline
%% \end {tabular}
%% }
%% \end{center}

%% If we model this function as $f: {\mathbb{Z}}_{8} \rightarrow
%% {\mathbb{Z}}_{8}$, then using the results of \cite{singmaster}
%% \cite{chen_95} \cite{chen_96}, we deduce that this function is not a
%% polynomial function over ${\mathbb{Z}}_{8}$. However, applying
%% Largange's interpolation formula over ${\mathbb{F}}_{2^3}$, we obtain
%% the following polynomial function representation, $Y =
%% (\alpha^2+1)A^4+(\alpha^2+1)A^2$, where $P(\alpha) = \alpha^3 +
%% \alpha + 1 = 0$. 
%% }
%% \end{Example}

An important property of Galois fields is that for all elements $A \in
{\mathbb{F}}_q, A^q = A$, and hence $A^q - A =  0.$ Therefore, the
polynomial $x^q -x$ {\it vanishes} on all points in
$\Fq$. Consequently, any polynomial $\F(X)$ can be reduced $\pmod{ X^q
  - X}$ to obtain a canonical representation $\F(X) \pmod{X^q - X}$
with degree at most $q-1$. 
%Such {\it vanishing polynomials} will form
%an important part of our formulation.

@@

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{../new_mmcircuit.eps}
	\end{center}
	\caption{{\it Montgomery} multiplication over $\mathbb{F}_{2^k}$
          using four Montgomery reductions.}
	\label{fig:mm4}
\end{figure}

>>>>

A Galois field is a field with a finite number of elements. The number
of elements $q$ of the field is a power of a prime integer --- i.e. $q
= p^k$, where $p$ is a prime integer, and $k \geq 1$ is a positive
integer \cite{galois_field:mceliece}. Galois fields are denoted as
${\mathbb{F}}_{q}$ and also $GF(q =p^k)$. We are interested in fields
where $p = 2$ and $k >1$ --- i.e. {\it   binary Galois extension
  fields} $\mathbb{F}_{2^k}$ --- as they are widely employed in
hardware implementations of cryptography primitives.

To construct ${\mathbb{F}}_{2^k}$, we take the polynomial ring
${\mathbb{F}}_2[x]$, where ${\mathbb{F}}_{2} = \{0, 1\}$, and an
irreducible  polynomial $P(x) \in {\mathbb{F}}_2[x]$ of degree $k$, and
construct ${\mathbb{F}}_{2^k}$ as ${\mathbb{F}}_2[x] \pmod{   P(x)}$. As a
result, all field operations are performed 
modulo the irreducible polynomial $P(x)$ and the coefficients are
reduced modulo $p=2$. %=2$; due to which $-1 = +1$ over $\Fkk$. 
Any element $A \in {\mathbb{F}}_{2^k}$ can be represented in polynomial
form as $A = a_0 +  a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}$, where
$a_i \in {\mathbb{F}}_2, i = 0, \dots, k-1$, and $\alpha$ is the root of
the irreducible polynomial, {\it i.e.} $P(\alpha)=0$. The field
$\Fkk$ can therefore be construed as a $k$-dimensional vector space
over ${\mathbb{F}}_2$.

@@

\begin{Example}\label{ex:1}
Let us construct ${\mathbb{F}}_{2^4}$ as ${\mathbb{F}}_2[x] \pmod{
  P(x)}$, where $P(x)=x^4+x^3+1 \in {\mathbb{F}}_2[x]$ is an
irreducible polynomial of degree $k =4$. Let $\alpha$ be the root of
$P(x)$, i.e. $P(\alpha)=0$.  Any element $A \in {\mathbb{F}}_2[x]
\pmod{ x^4 + x^3 + 1}$ has a representation of the type: $A = a_3 x^3
+ a_2 x^2 +  a_1 x + a_0$ where the coefficients $a_3, \dots, a_0$ are
in ${\mathbb{F}}_2 = \{0, 1\}$. Since there are only 16 such
polynomials, we obtain 16 elements in the field
${\mathbb{F}}_{16}$. Each element in can then be viewed as a 4-bit
vector over ${\mathbb{F}}_2$: ${\mathbb{F}}_{16}=\{(0000),(0001),
\dots (1110),(1111)\}$.  Each element also has an exponential
representation; all three representations are shown in Table
\ref{tab:gfelement}. For example, consider the element $\alpha^{12}$.
Computing $\alpha^{12} \pmod{ \alpha^4+\alpha^3+1} = \alpha + 1
= (0011)$; hence we have the three equivalent representations.

\vspace{-0.1in}
\begin{table}[h]
\begin{center}
{\small
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
\pmod{x^4+x^3+1}$}\label{tab:gfelement}  
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
}
\end{center}
\end{table}
\end{Example}

%%%%%%%%%
{\bf Polynomial Functions $f: \Fkk \rightarrow \Fkk$:} 
Arbitrary mappings among $k$-bit vectors can be constructed; each such
mapping generates a function $f: \B^k \rightarrow \B^k$. Since every
$k$-bit vector can be construed as an element in $\Fkk$ (as shown in
the above example), every such function corresponds to a function over
Galois fields: $f: \Fkk \rightarrow \Fkk$. Every such function is also
a polynomial function.

\begin{Theorem}
From \cite{ff:1997}: 
Let $\Fq$ be a Galois field of $q$ elements where $q$ is a power of a
prime integer. Any  function $f: \Fq \to \Fq$ is a polynomial function
over $\Fq$, that is there exists a polynomial $\F \in \Fq [x]$ such
  that $f(a) = \F(a)$, for all $a \in \Fq$. 
\end{Theorem}

@@

\begin{Example} {\it
Let $A = \{a_2, a_1, a_0\}$ and $Y = \{y_2, y_1,
y_0\}$ be 3-bit vectors.  Consider the function $Y[2:0] = A[2:0]
>> 1$, i.e. a {\bf bit-vector right shift} operation on $A$. 
The function maps as follows:

\begin{center}
{\small
\begin{tabular}{c|ccc|c||c|ccc|c}
$\{a_2a_1a_0\}$  & $A$ &$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$ &
  $\{a_2a_1a_0\}$  &$A$&$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$\\
\hline
000  &0 &$\rightarrow$& 000 & 0 & 100  &$\alpha^2$ &$\rightarrow$& 010 &  $\alpha$ \\
001  &1 &$\rightarrow$& 000 & 0 & 101  &$\alpha^2 + 1$ &$\rightarrow$&010 & $\alpha$ \\
010  &$\alpha$ & $\rightarrow$ & 001& 1 & 110  &$\alpha^2 + \alpha$&$\rightarrow$& 011 &$\alpha + 1$ \\   
011  &$\alpha + 1$ &$\rightarrow$& 001 &1 & 111& $\alpha^2 + \alpha + 1$ &$\rightarrow$& 011 &$\alpha + 1$\\
\hline
\end {tabular}
}
\end{center}

If we model this function as $f: {\mathbb{Z}}_{8} \rightarrow
{\mathbb{Z}}_{8}$, then using the results of \cite{singmaster}
\cite{chen_95} \cite{chen_96}, we deduce that this function is not a
polynomial function over ${\mathbb{Z}}_{8}$. However, applying
Largange's interpolation formula over ${\mathbb{F}}_{2^3}$, we obtain
the following polynomial function representation, $Y =
(\alpha^2+1)A^4+(\alpha^2+1)A^2$, where $P(\alpha) = \alpha^3 +
\alpha + 1 = 0$. 
}
\end{Example}

An important property of Galois fields is that for all elements $A \in
{\mathbb{F}}_q, A^q = A$, and hence $A^q - A =  0.$ Therefore, the
polynomial $x^q -x$ {\it vanishes} on all points in
$\Fq$. Consequently, any polynomial $\F(X)$ can be reduced $\pmod{ X^q
  - X}$ to obtain a canonical representation $\F(X) \pmod{X^q - X}$
with degree at most $q-1$. Such {\it vanishing polynomials} will form
an important part of our ideal membership formulation.

@@

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{new_mmcircuit.eps}
	\end{center}
	\caption{{\it Montgomery} multiplication over $\mathbb{F}_{2^k}$
          using four Montgomery reductions.}
	\label{fig:mm4}
\end{figure}
